---
title: "new hk macro mod"
output: html_document
date: "2023-06-29"
editor_options: 
  chunk_output_type: console
---
```{r}
#load packages
library(pacman)

p_load(tidyverse, stargazer, vars, forecast, bvartools, readxl, tseries, lubridate, RColorBrewer, ggfortify, Metrics, BVAR, forestmangr, tsDyn, x12)

#redefine select function
select<-dplyr::select

```

```{r}
df<-read.csv("clean data2.csv")

worldgdp_df<-read.csv('world gdp.csv')
worldgdp_df$date<-mdy(worldgdp_df$date)
worldgdp_df<-worldgdp_df%>%
  na.omit()%>%
  mutate(worldgdp=jp_gdp+ gm_gdp+ fr_gdp+ uk_gdp+ us_gdp)%>%
  filter(date<mdy('04-01-2023'))

rates_df<-read.csv('rates.csv')
rates_df$date<-mdy(rates_df$date)

unitvalueim_df<-read.csv('Unit Value Index Imports.csv')
unitvalueim_df$date<-mdy(unitvalueim_df$date)
```

turn date from character to date 
```{r}
df$date<-mdy(df$date)
```

###################################################################
data for new var model
```{r}


endo_df<-df%>%
  mutate(lrgdp=log(hk_rgdp),
         lrgdp_d=lrgdp-lag(lag(lag(lag(lrgdp)))),
         lcpi=log(hk_cpi),
         lcpi_d=lcpi-lag(lag(lag(lag(lcpi)))))%>%
  select(c('lrgdp_d','date','lcpi_d'))



inv_data<-read.csv('investment data.csv')

inv_data$date<-mdy(inv_data$date)

temp<-inv_data%>%
  mutate(linv=log(hk_gfcf),
         linv_d=linv-lag(linv))%>%
  select(date, linv_d)

endo_df<-left_join(endo_df, temp)

exo_df<-worldgdp_df%>%
  mutate(lwgdp=log(worldgdp),
         lwgdp_d=lwgdp-lag(lwgdp))%>%
  select(date, lwgdp_d)

temp<-rates_df%>%
  select(date, hk_blr)

exo_df<-left_join(exo_df, temp)

unitvalueim_df$hk_uvi<-as.numeric(unitvalueim_df$hk_uvi)

#yoy since uvi is monthly data
temp<-unitvalueim_df%>%
  mutate(luvi=log(hk_uvi),
         luvi_d=luvi-lag(lag(lag(lag(lag(lag(lag(lag(lag(lag(lag(lag(luvi)))))))))))))%>%
  select(date, luvi_d)


exo_df<-left_join(exo_df, temp)

sars<-data.frame(date=exo_df$date, sars=0)
sars$sars[sars$date>=mdy('01-01-2002') & sars$date<=mdy('07-01-2003')]<-1



exo_df<-left_join(exo_df, sars)

exo_df$hk_blr<-as.numeric(exo_df$hk_blr)

endo_df<-endo_df%>% na.omit()%>%filter(date<mdy('01-01-2020'))
endo_df<-endo_df%>% na.omit()%>%filter(date>=mdy('06-01-1991'))
exo_df<-exo_df%>%na.omit()%>%filter(date<mdy('01-01-2020'))


```



add lag to exo 2 data (sars hk_blr)
```{r}
exo2_df<-data.frame(sars=exo_df$sars,
                    hk_blr=exo_df$hk_blr,
                    date=exo_df$date)

exo2_df<-exo2_df%>%
  mutate(sars_l1=lag(sars),
         sars_l2=lag(sars_l1), 
         sars_l3=lag(sars_l2),
         hk_blr_l1=lag(hk_blr),
         hk_blr_l2=lag(hk_blr_l1),
         hk_blr_l3=lag(hk_blr_l2))
```

df with endogenous variables: endo_df
df with exogenous variables: exo_df


variables in endo_df:
lrgdp_d
lcpi_d  
linv_d

variables in exo_df:
lwgdp_d 
hk_blr  
luvi_d  
sars   

(time from 1992 q1 to 2019 q4)
make time series objects
```{r}
#endogenous variables
lrgdp_d_ts<-ts(endo_df$lrgdp_d, start=c(1991,2),end=c(2019,4),4)
lcpi_d_ts<-ts(endo_df$lcpi_d  , start=c(1991,2),end=c(2019,4),4)
linv_d_ts<-ts(endo_df$linv_d, start=c(1991,2),end=c(2019,4),4)


#exogenous variables
lwgdp_d_ts<-ts(exo_df$lwgdp_d , start=c(1991,2),end=c(2019,4),4)
luvi_d_ts<-ts(exo_df$luvi_d , start=c(1991,2),end=c(2019,4),4)

#exogenous 2 variables
sars_ts_l1<-ts(exo2_df$sars_l1 , start=c(1991,2),end=c(2019,4),4)
sars_ts_l2<-ts(exo2_df$sars_l2 , start=c(1991,2),end=c(2019,4),4)
sars_ts_l3<-ts(exo2_df$sars_l3 , start=c(1991,2),end=c(2019,4),4)
hk_blr_ts_l1<-ts(exo2_df$hk_blr_l1 , start=c(1991,2),end=c(2019,4),4)
hk_blr_ts_l2<-ts(exo2_df$hk_blr_l2 , start=c(1991,2),end=c(2019,4),4)
hk_blr_ts_l3<-ts(exo2_df$hk_blr_l3 , start=c(1991,2),end=c(2019,4),4)

```


TRYING OUT X12 SEASONALITY ADJUSTMENT
```{r}
#hk_rgdp_ts<-ts(df$hk_rgdp[!is.na(df$hk_rgdp)], start=c(1973,1), frequency = 4)
#x12out <- x12(hk_rgdp_ts,transform="auto",
#		arima=c(0,1,1),sarima=c(0,1,1),regvariables="lpyear",
#		sigmalim=c(2.0,3.0),outlier="all",critical=list(LS=3.5,TC=3),
#		seasonalma="s3x3")

#summary(x12(hk_rgdp_ts,arima=c(0,1,1),sarima=c(0,1,1)))
```







merge data to for var
training period (1992 q1- 2016 q4)
testing period (2017 q1- 2019 q4)
```{r}
endo_dat<-window(ts.union(lcpi_d_ts,lrgdp_d_ts,linv_d_ts), start=c(1991,2), end=c(2016,4))
exo_dat<-window(ts.union(lwgdp_d_ts,luvi_d_ts), start=c(1991,2), end=c(2016,4))
exo2_dat<-window(ts.union(sars_ts_l1, sars_ts_l2, sars_ts_l3, hk_blr_ts_l1, hk_blr_ts_l2, hk_blr_ts_l3), start=c(1991,2), end=c(2016,4))

```





```{r}
#set dates
q_ahead<-12
start_date<-ymd("2016-12-01")
test_date<-data.frame(date=ymd("1999-10-11")) #initialize date 
test_date[1,1]<-start_date %m+% months(3)

for (i in 2:q_ahead)
{
  test_date[i,1]<-test_date[i-1,1] %m+% months(3)
}

```



run var for exo
```{r}
exo_mod<-VAR(exo_dat,p=3, type='const',season=4,  exogen=exo2_dat, ic='AIC')


sars<-1:q_ahead
sars[1:q_ahead]<-0

#make blr constant =5
blr<-1:q_ahead
blr[1:q_ahead]<-5

hk_blr_ts_l1<-ts(blr, start=c(2017,1), frequency = 4)
hk_blr_ts_l2<-ts(blr, start=c(2017,1), frequency = 4)
hk_blr_ts_l3<-ts(blr, start=c(2017,1), frequency = 4)
sars_ts_l1<-ts(sars, start=c(2017,1), frequency = 4)
sars_ts_l2<-ts(sars, start=c(2017,1), frequency = 4)
sars_ts_l3<-ts(sars, start=c(2017,1), frequency = 4)

exo_dv<-window(ts.union(sars_ts_l1,sars_ts_l2, sars_ts_l3, hk_blr_ts_l1, hk_blr_ts_l2, hk_blr_ts_l3), start=c(2017,1))

exo_pred<-predict(exo_mod, n.ahead=q_ahead, dumvar=exo_dv)
```


make exogen ts data for full model
```{r}
#add lag to variables in exo_df
exo_df<-exo_df%>%
  mutate(lwgdp_d_l1=lag(lwgdp_d),
         lwgdp_d_l2=lag(lwgdp_d_l1),
         lwgdp_d_l3=lag(lwgdp_d_l2),
         luvi_d_l1=lag(luvi_d),
         luvi_d_l2=lag(luvi_d_l1),
         luvi_d_l3=lag(luvi_d_l2))

lwgdp_d_ts<-ts(exo_df$lwgdp_d, start=c(1991,2), frequency = 4)
lwgdp_d_ts_l1<-ts(exo_df$lwgdp_d_l1, start=c(1991,2), frequency = 4)
lwgdp_d_ts_l2<-ts(exo_df$lwgdp_d_l2, start=c(1991,2), frequency = 4)
lwgdp_d_ts_l3<-ts(exo_df$lwgdp_d_l3, start=c(1991,2), frequency = 4)

luvi_d_ts<-ts(exo_df$luvi_d, start=c(1991,2), frequency = 4)
luvi_d_ts_l1<-ts(exo_df$luvi_d_l1, start=c(1991,2), frequency = 4)
luvi_d_ts_l2<-ts(exo_df$luvi_d_l2, start=c(1991,2), frequency = 4)
luvi_d_ts_l3<-ts(exo_df$luvi_d_l3, start=c(1991,2), frequency = 4)

sars_ts<-ts(exo2_df$sars, start=c(1991,2), frequency=4)
sars_ts_l1<-ts(exo2_df$sars_l1, start=c(1991,2), frequency=4)
sars_ts_l2<-ts(exo2_df$sars_l2, start=c(1991,2), frequency=4)
sars_ts_l3<-ts(exo2_df$sars_l3, start=c(1991,2), frequency=4)

hk_blr_ts<-ts(exo2_df$hk_blr, start=c(1991,2), frequency=4)
hk_blr_ts_l1<-ts(exo2_df$hk_blr_l1, start=c(1991,2), frequency=4)
hk_blr_ts_l2<-ts(exo2_df$hk_blr_l2, start=c(1991,2), frequency=4)
hk_blr_ts_l3<-ts(exo2_df$hk_blr_l3, start=c(1991,2), frequency=4)

full_dat<-window(ts.union(lwgdp_d_ts, luvi_d_ts, hk_blr_ts, sars_ts, lcpi_d_ts, linv_d_ts, lrgdp_d_ts), start=c(1991,2), end=c(2016,4))


exo_dat<-window(ts.union(lwgdp_d_ts_l1, lwgdp_d_ts_l2, lwgdp_d_ts_l3,luvi_d_ts_l1, luvi_d_ts_l2, luvi_d_ts_l3,sars_ts_l1,sars_ts_l2, sars_ts_l3, hk_blr_ts_l1, hk_blr_ts_l2, hk_blr_ts_l3), start=c(1991,2), end=c(2016,4))


full_mod<-VAR(endo_dat, p=3, season=4, type='const', exogen= exo_dat, ic='AIC')

summary(full_mod)

stargazer(full_mod$varresult$lrgdp_d_ts)



```


make prediction with full_mod model
```{r}
exo_pred<-data.frame(lwgdp_d_p=exo_pred$fcst$lwgdp_d_ts[,1],
                     luvi_d_p=exo_pred$fcst$luvi_d_ts[,1],
                     hk_blr_p=blr,
                     sars=sars,
                     date=test_date)

exo_pred_act<-exo_df%>%
  filter(date<test_date[1,1])

exo_pred_act<-rbind(left_join(exo_df, exo_pred), right_join(exo_df, exo_pred))




for(i in 1:sum(is.na(exo_pred_act$lwgdp_d_p)))
{
  exo_pred_act$lwgdp_d_p[i]<-exo_pred_act$lwgdp_d[i]
  exo_pred_act$luvi_d_p[i]<-exo_pred_act$luvi_d[i]
  exo_pred_act$hk_blr_p[i]<-exo_pred_act$hk_blr[i]
}

exo_pred_act<-exo_pred_act%>%
  mutate(lwgdp_d_p_l1=lag(lwgdp_d_p),
         lwgdp_d_p_l2=lag(lwgdp_d_p_l1),
         lwgdp_d_p_l3=lag(lwgdp_d_p_l2),
         luvi_d_p_l1=lag(luvi_d),
         luvi_d_p_l2=lag(luvi_d_l1),
         luvi_d_p_l3=lag(luvi_d_l2),
         hk_blr_p_l1=lag(hk_blr_p),
         hk_blr_p_l2=lag(hk_blr_p_l1),
         hk_blr_p_l3=lag(hk_blr_p_l2))

temp<-exo_pred_act%>%
    select(lwgdp_d_p_l1, lwgdp_d_p_l2, lwgdp_d_p_l3, luvi_d_p_l1, luvi_d_p_l2, luvi_d_p_l3, hk_blr_p_l1, hk_blr_p_l2, hk_blr_p_l3, sars, date)%>%
  filter(date>=test_date[1,1])





lwgdp_d_ts_l1<-ts(temp$lwgdp_d_p_l1, start=c(2017,1), frequency = 4)
lwgdp_d_ts_l2<-ts(temp$lwgdp_d_p_l2, start=c(2017,1), frequency = 4)
lwgdp_d_ts_l3<-ts(temp$lwgdp_d_p_l3, start=c(2017,1), frequency = 4)

luvi_d_ts_l1<-ts(temp$luvi_d_p_l1, start=c(2017,1), frequency = 4)
luvi_d_ts_l2<-ts(temp$luvi_d_p_l2, start=c(2017,1), frequency = 4)
luvi_d_ts_l3<-ts(temp$luvi_d_p_l3, start=c(2017,1), frequency = 4)

hk_blr_ts_l1<-ts(temp$hk_blr_p_l1, start=c(2017,1), frequency = 4)
hk_blr_ts_l2<-ts(temp$hk_blr_p_l2, start=c(2017,1), frequency = 4)
hk_blr_ts_l3<-ts(temp$hk_blr_p_l3, start=c(2017,1), frequency = 4)

sars_ts_l1<-ts(temp$sars, start=c(2017,1), frequency=4)
sars_ts_l2<-ts(temp$sars, start=c(2017,1), frequency=4)
sars_ts_l2<-ts(temp$sars, start=c(2017,1), frequency=4)


exo_dv<-window(ts.union(lwgdp_d_ts_l1,lwgdp_d_ts_l2,lwgdp_d_ts_l3, luvi_d_ts_l1,luvi_d_ts_l2,luvi_d_ts_l3,sars_ts_l1,sars_ts_l2, sars_ts_l3, hk_blr_ts_l1, hk_blr_ts_l2, hk_blr_ts_l3), start=c(2017,1), end=(c(2019,4)))




endo_pred<-predict(full_mod,n.ahead=q_ahead, dumvar=exo_dv)


endo_pred<-data.frame(lrgdp_d_p=endo_pred$fcst$lrgdp_d_ts[,1],
                      lcpi_d_p=endo_pred$fcst$lcpi_d_ts[,1],
                      linv_d_p=endo_pred$fcst$linv_d_ts[,1],
                      date=test_date)



#extract predicted for exo and endo
endo_pred_act<-left_join(endo_df, endo_pred)

```

plot predictions
```{r}
for(i in 4:103){
  endo_pred_act$lrgdp_d_p[i]<-as.data.frame(fitted(full_mod))$lrgdp_d_ts[i-3]
  endo_pred_act$lcpi_d_p[i]<-as.data.frame(fitted(full_mod))$lcpi_d_ts[i-3]
  endo_pred_act$linv_d_p[i]<-as.data.frame(fitted(full_mod))$linv_d_ts[i-3]
}

as.data.frame(fitted(full_mod))$lrgdp_d_ts

#plot lrgdp_d with pred
endo_pred_act%>%
  pivot_longer(c(lrgdp_d, lrgdp_d_p))%>%
  select(c(value, date, name))%>%
  na.omit()%>%
  ggplot(aes(x=date, y=value, color=name))+
  geom_line()+
  theme_bw()+
  geom_vline(xintercept=ymd('2017-01-01'))

#plot lcpi_d with pred
endo_pred_act%>%
  pivot_longer(c(lcpi_d, lcpi_d_p))%>%
  select(c(value, date, name))%>%
  na.omit()%>%
  ggplot(aes(x=date, y=value, color=name))+
  geom_line()+
  theme_bw()+
  geom_vline(xintercept=ymd('2017-01-01'))

#plot linv_d with pred
endo_pred_act%>%
  pivot_longer(c(linv_d, linv_d_p))%>%
  select(c(value, date, name))%>%
  na.omit()%>%
  ggplot(aes(x=date, y=value, color=name))+
  geom_line()+
  theme_bw()+
  geom_vline(xintercept=ymd('2017-01-01'))

#plot lwgdp_d with pred
exo_pred_act%>%
  pivot_longer(c(lwgdp_d, lwgdp_d_p))%>%
  select(c(value, date, name))%>%
  na.omit()%>%
  ggplot(aes(x=date, y=value, color=name))+
  geom_line()+
  theme_bw()+
  geom_vline(xintercept=ymd('2017-01-01'))

#plot luvi_d with pred
exo_pred_act%>%
  pivot_longer(c(luvi_d, luvi_d_p))%>%
  select(c(value, date, name))%>%
  na.omit()%>%
  ggplot(aes(x=date, y=value, color=name))+
  geom_line()+
  theme_bw()+
  geom_vline(xintercept=ymd('2017-01-01'))
```



rmse for endogenous variable predictions
```{r}
rmse(endo_pred_act$lrgdp_d[endo_pred_act$date>=ymd('2017-1-1')], endo_pred_act$lrgdp_d_p[endo_pred_act$date>=ymd('2017-1-1')])

rmse(endo_pred_act$lcpi_d[endo_pred_act$date>=ymd('2017-1-1')], endo_pred_act$lcpi_d_p[endo_pred_act$date>=ymd('2017-1-1')])

rmse(endo_pred_act$linv_d[endo_pred_act$date>=ymd('2017-1-1')], endo_pred_act$linv_d_p[endo_pred_act$date>=ymd('2017-1-1')])

rmse(endo_pred_act$lrgdp_d[endo_pred_act$date>=ymd('1992-1-1')], endo_pred_act$lrgdp_d_p[endo_pred_act$date>=ymd('1992-1-1')])

rmse(endo_pred_act$lcpi_d[endo_pred_act$date>=ymd('1992-1-1')], endo_pred_act$lcpi_d_p[endo_pred_act$date>=ymd('1992-1-1')])

rmse(endo_pred_act$linv_d[endo_pred_act$date>=ymd('1992-1-1')], endo_pred_act$linv_d_p[endo_pred_act$date>=ymd('1992-1-1')])


```


predicted and actual rgdp
```{r}
temp<-df%>%
  select(date, hk_rgdp)


endo_pred_act<-left_join(endo_pred_act, temp)%>%
  mutate(lrgdp_p=log(hk_rgdp))

for(i in (sum(is.na(endo_pred_act$lrgdp_d_p))+1):(sum(is.na(endo_pred_act$lrgdp_d_p))+q_ahead))
{
  endo_pred_act$lrgdp_p[i]<-endo_pred_act$lrgdp_p[i-1]+endo_pred_act$lrgdp_d_p[i]
}

for(i in (1):((length(endo_pred_act$lrgdp_d_p))-q_ahead-1))
{
  endo_pred_act$lrgdp_p[i]<-NA
}

endo_pred_act%>%
  mutate(rgdp_p=exp(lrgdp_p))%>%
  pivot_longer(c('rgdp_p','hk_rgdp'))%>%
  ggplot(aes(x=date, y=value, color=name))+
  geom_line()+
  theme_bw()+
  geom_vline(xintercept=mdy('01-01-2017'))+
  ylab('Real GDP')+
  ggtitle('Actual and modeled real gdp')
```


predicted and actual cpi
```{r}
temp<-df%>%
  select(date, hk_cpi)

endo_pred_act<-left_join(endo_pred_act, temp)%>%
  mutate(lcpi_p=log(hk_cpi))

for(i in (sum(is.na(endo_pred_act$lcpi_d_p))+1):(sum(is.na(endo_pred_act$lcpi_d_p))+q_ahead))
{
  endo_pred_act$lcpi_p[i]<-endo_pred_act$lcpi_p[i-1]+endo_pred_act$lcpi_d_p[i]
}

for(i in (1):((length(endo_pred_act$lcpi_d_p))-q_ahead-1))
{
  endo_pred_act$lcpi_p[i]<-NA
}

endo_pred_act%>%
  mutate(cpi_p=exp(lcpi_p))%>%
  pivot_longer(c('cpi_p','hk_cpi'))%>%
  ggplot(aes(x=date, y=value, color=name))+
  geom_line()+
  theme_bw()+
  geom_vline(xintercept=mdy('01-01-2017'))+
  ylab('Consumer price index')+
  ggtitle('Actual and modeled CPI')

```


predicted and actual investment
```{r}
temp<-inv_data%>%
  select(date, hk_gfcf)

endo_pred_act<-left_join(endo_pred_act, temp)%>%
  mutate(linv_p=log(hk_gfcf))

for(i in (sum(is.na(endo_pred_act$linv_d_p))+1):(sum(is.na(endo_pred_act$linv_d_p))+q_ahead))
{
  endo_pred_act$linv_p[i]<-endo_pred_act$linv_p[i-1]+endo_pred_act$linv_d_p[i]
}

for(i in (1):((length(endo_pred_act$linv_d_p))-q_ahead-1))
{
  endo_pred_act$linv_p[i]<-NA
}


endo_pred_act%>%
  mutate(inv_p=exp(linv_p))%>%
  pivot_longer(c('inv_p','hk_gfcf'))%>%
  ggplot(aes(x=date, y=value, color=name))+
  geom_line()+
  theme_bw()+
  geom_vline(xintercept=mdy('01-01-2017'))+
  ylab('Gross fixed capital formation (usd millions)')+
  ggtitle('Actual and modeled domestic investments')


data.frame

class(lcpi_d_ts)

```


```{r}
full_mod2<-VAR(full_dat, p=3, season=4, type='const', ic='AIC')



plot(irf(full_mod2, impulse = 'linv_d_ts', response= 'lrgdp_d_ts'))
plot(irf(full_mod2, impulse = 'lcpi_d_ts', response= 'lrgdp_d_ts'))
plot(irf(full_mod2, impulse = 'lwgdp_d_ts', response= 'lrgdp_d_ts'))
plot(irf(full_mod2, impulse = 'hk_blr_ts', response= 'lrgdp_d_ts'))
plot(irf(full_mod2, impulse = 'luvi_d_ts', response= 'lrgdp_d_ts'))
plot(irf(full_mod2, impulse = 'sars_ts', response= 'lrgdp_d_ts'))


```









